---
title: "General Social Survey"
author: "Kaylie Fagoaga"
date: "Summer 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: CAPPUN is independent of OWNGUN 

$H_A$: CAPPUN id dependent of OWNGUN

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
CAPPUN is a categorical variable,  
OWNGUN is a categorical variable,   
so I use the cat~cat helper file

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")

```

This graph shows that people who own a gun are more likely to favor capital punishment.     




<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~CAPPUN + OWNGUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

The numerical summary shows that 75% of gun owners favor CAPPUN. 25% of gun owners oppose CAPPUN.   


### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This p-value for the Chi-squared test is p=0.02002.   
The p-value for fisher exact test was p=0.01651,which is also less than 0.05. 

The odds ratio says that gun owners are 2.271 times more likely to favor capital punishment than those who don't own a gun.  

This means we reject the null hypothesis and conclude that the opinion on capital punishment (CAPPUN) is dependent on gun ownership (OWNGUN) because both p-value's are less than 0.05.    



<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->
Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is a difference of Number of education of those who did/didn't vote in 2012

$H_A$: There is no difference of Number of education of those who did/didn't vote in 2012

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

EDUC is numerical variable  
VOTE12 is categorical variable  
So I use the num~cat helper file

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results
```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

The boxplot shows that people who voted in 2012 have a higher education than people who did not vote in 2012.


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results
```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```

The table shows that the mean of years of education for people who voted in 2012 is 15.7 whereas the mean of of years education for people who didn't vote is 13.4.


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results
```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

The t-test shows that the p-value is 0.0000001248.  
This means we reject the null hypothesis that education (EDUC) is independent of those who did/didn't vote in 2012 (VOTE12) because the p-value is smaller than 0.05.   
Between those who did/didn't vote in 2012, we conclude that it is dependent on the number of years of education.   

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a person's performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$: Person's score on the wordsum test (WORDSUM) is independent of level of schooling (DEGREE)  

$H_A$: Person's score on the wordsum test (WORDSUM) is dependent of level of schooling (DEGREE)   

## Methods
WORDSUM is a numerical variable  
DEGREE is categorical variable  
So I use the Num~cat helper file  

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
plot(WORDSUM~DEGREE,data=GSSdata)
```

This boxplot shows that people who are Graduates have a tendency of having a higher score on their wordsum test (WORDSUM) than people with fewer years of education.  


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

This numerical summary shows the mean of Wordsum test scores between the years of education.    
It shows that Graduates have a mean of 9.5 which is the highest among the degrees.   

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results
```{r}
anova=aov(WORDSUM~DEGREE,data= GSSdata)
summary(anova)
```  

This shows that we reject the null hypothesis because the p-value is 0.0000002 which is less than 0.05.


```{r}
par(mar = c(7, 8.7, 2, 2))
plot(TukeyHSD(anova),las=2)
```

The tukey test shows that there is no difference between high school and less than high school.  
There is a difference between junior c ollege and high school.  
There is a difference between bachelors degree and high school.   
No difference between high school and graduate.    
No difference between junior college and graduate.  
No difference between bachelors and graduates.  

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

#### Question 1: Is gun onwership independent of opinion on capital punishment? 
The summary result and the inferential results show that they are dependent. We conclude that the opinion on capital punishment is dependent on gun ownership

#### Question 2: Is there a difference in years of education between the those who did or did not vote in 2012?
The summary result and inferential results show that they are not independent. Between those who did/didn't vote in 2012, we conclude that it is dependent on the number of years of education.

#### Question 3: Is a persons performance on the wordsum test independent of their level of schooling?
The summary and inferential results show that they are dependent. People who are Graduates have a tendency of having a higher score on their wordsum test than people with fewer years of education.